### 2010-03-30 papers/icfp-2010-desc/paper_discussion.tex
Some suggestions to fix some typos

As presented, our type theory suffers from a major weakness. Indeed,
we are subject to Girard's paradox, as we assume that $\Set$ lives in
$\Set$. We made that choice for presentational convenience, as
universe stratification is orthogonal to our work. Nonetheless, our
universes of description rather naturally leads itself to
stratification. Unsurprisingly, $\IDesc{\!}$ at level $l$ is of type
$\Set^{\blue{l+1}}$. Similarly, the interpretation of $\IDesc{\!}$ at
level $l$ is an object of type $\Set^{\blue{l}}$:

@@

In particular, our approach is similar in spirit with polytypic
programming, as initiated by PolyP~\cite{jansson:polyp}. Indeed,
generic functions, in our system, are built by induction on the
pattern functor. Unlike PolyP, we do not have to resort to
pre-processing code: our data-types are, natively, nothing but codes.

Our approach also support the Generic
Haskell~\cite{hinze:generic-haskell} model. This model, based on
type-indexed data types, corresponds to computing new data-types from
others. This is natural in our system, as data-types descriptions are
first-class.

Interestingly, our generic induction principle allows a programming
style reminiscent to the Scrap Your Boilerplate~\cite{spj:syb} (SYB)
approach. Like SYB, we provide a generic operator to manipulate and
carry operations over data-types. However, unlike SYB, we do not need
support from the compiler, as introduced by the $\CN{Typeable}$
class. Indeed, in our setting, data-types are reflected in the
language, hence our ability to manipulate them directly.

@@

The interest in generic programming for dependent types is not new
either. Norell~\cite{norell:msc-thesis} have shown the benefit of
polytypic programming in the setting of Alfa, a precursor of
Agda. Similarly, Verbruggen et al.~\cite{verbruggen:polytype-prog-coq,
  verbruggen:polytype-coq} have developed a framework for polytypic
programming in the Coq theorem prover. However, these works aim at
\emph{modelling} Generic Haskell in a dependently-typed setting, for
the purpose of proving correctness properties of Haskell code. Our
approach is different in that we aim at building a foundation for
data-types in a dependently typed theory.

Closer to us is the work by Benke et
al.~\cite{benke:universe-generic-prog}. This seminal work introduced
the usage of universes for developing generic programs. Moreover, our
own universes are rather similar to theirs: our universe of
descriptions is similar to their universe of iterated induction, and
our universe of indexed descriptions is isomorphic to their universe of
finitary indexed induction. This is not surprising, as we share the
same source of inspiration, namely induction recursion.

However, we differ in several ways. First, they adopt a generative
perspective: each universe extends the base type theory with both type
formers and elimination rules. Thanks to the levitation, we only rely
on a generic induction and a specialised
$\switchD{\!}{\!}{\!}$. Second, the authors develop several universes,
each with its own strengths and weaknesses, but lack a formal,
unifying universe. With indexed descriptions, we subsume simple
descriptions, hence avoiding the move to and from several
codings. Finally, the authors often resort to an extensional notion of
equality, while we have given an equality-agnostic
presentation. Moreover, our presentation is arranged so as to use
definitional equality as much as possible. Hence, in practice, the
developer is relieved from many trivial proof obligations.

>>>>

As presented, our type theory suffers from a major weakness. Indeed,
we are subject to Girard's paradox, as we assume that $\Set$ lives in
$\Set$. We made that choice for presentational convenience, as
universe stratification is orthogonal to our work. Nonetheless, our
universe of description rather naturally leads itself to
stratification. Unsurprisingly, $\IDesc{\!}$ at level $l$ is of type
$\Set^{\blue{l+1}}$. Similarly, the interpretation of $\IDesc{\!}$ at
level $l$ is an object of type $\Set^{\blue{l}}$:

@@

In particular, our approach is similar in spirit with polytypic
programming, as initiated by PolyP~\cite{jansson:polyp}. Indeed,
generic functions, in our system, are built by induction on the
pattern functor. Unlike PolyP, we do not have to resort to pre-process
data-type definitions: our data-types are, natively, nothing but
codes.

Our approach also support the Generic
Haskell~\cite{hinze:generic-haskell} model. This model, based on
type-indexed data types, enables computing new data-types from
others. This is natural in our system, as data-types descriptions are
first-class.

Another generic programming framework is Scrap Your
Boilerplate~\cite{spj:syb} (SYB). Our proposal is different in various
ways. The corner stone of SYB is the \emph{spine} view of data-type
constructors. A data-type is a spine composed by a constructor applied
to some arguments. Further, this spine is equipped with some
combinators including, primarily, an iterator. In this setting,
generic programs are written by composing these combinators. This
relies on a $\CN{Typeable}$ type-class, allowing dynamic dispatch to
data-type specific operations. As a result, SYB is not reflexive: it
is restricted to data-types instanciating $\CN{Typeable}$. Moreover,
it forbids type-indexed data types: it can only define generic
functions.

@@

The interest in generic programming for dependent types is not new
either. Norell~\cite{norell:msc-thesis} have shown the benefit of
polytypic programming in the setting of Alfa, a precursor of
Agda. Similarly, Verbruggen et al.~\cite{verbruggen:polytype-prog-coq,
  verbruggen:polytype-coq} have developed a framework for polytypic
programming in the Coq theorem prover. However, these works aim at
\emph{modelling} PolyP or Generic Haskell in a dependently-typed
setting, for the purpose of proving correctness properties of Haskell
code. Our approach is different in that we aim at building a
foundation for data-types, in a dependently-typed theory, for a
dependently-typed system.

Closer to us is the work by Benke et
al.~\cite{benke:universe-generic-prog}. This seminal work introduced
the usage of universes for developing generic programs. Our own
universes are rather similar to theirs: our universe of descriptions
is similar to their universe of iterated induction, and our universe
of indexed descriptions is isomorphic to their universe of finitary
indexed induction. This is not surprising, as we share the same source
of inspiration, namely induction recursion.

However, we differ in several ways. First, they adopt a generative
perspective: each universe extends the base type theory with both type
formers and elimination rules. Thanks to the levitation, we only rely
on a generic induction and a specialised
$\switchD{\!}{\!}{\!}$. Second, the authors do not tackle the issue of
\emph{programming} with such codes: it is not obvious, to us, how one
can define, manipulate, and program over these coded data-types. With
descriptions, we have shown how to abstract away codes and present a
convenient and familiar presentation to the developer. Finally, the
authors often resort to an extensional notion of equality, while we
have given an equality-agnostic presentation. Beside, our presentation
is arranged so as to use definitional equality as much as
possible. Hence, in practice, the developer is relieved from many
proof obligations.
